---
title: "10.introduction_to_dataset.Rmd"
author: "Guillaume Papuga"
date: "2023-11-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load data, echo = FALSE}
### Load the data
# Sorted dataset
data_plant_med = read.csv(here::here("data", "raw", "data_plant_med.csv"), 
                           header = T, row.names = 1) %>%
  mutate (date = as.Date(eventdate)) %>% # Convert into date
  select (-eventdate) # delete the original column

# Raw dataset


```



# Global analysis of the dataset

## Table 1 : general analysis

The aim is to built a table that summarise the general information contained in the dataset. 

```{r global analysis}

table(data_plant_med%>%filter(datasource == "inat")%>%select(species))

nrow(unique (data_plant_med%>%filter(datasource == "inat")%>%select(species)))


```

## Table 2 : species missing

```{r species missing}



```



# Temporal trends in the two datasets

## Number of observations 

```{r observation}
# Build the dataset
cumul_occ = data_plant_med %>%
  arrange (datasource, date) %>%
  group_by(datasource, date) %>%
  summarise(occ_date = n()) %>%
  mutate(cumul = cumsum(occ_date))

# Plot
ggplot(cumul_occ, aes(x = date, y = cumul, colour = datasource)) +
  geom_line() +
  theme_minimal()
```



## Trends in species richness

```{r species richness}
# Build the dataset
cumul_sp = data_plant_med %>%
  arrange (datasource, date) %>%
  distinct(species, datasource, .keep_all = TRUE) %>%
  group_by(datasource, date) %>%
  summarise(occ_date = n()) %>%
  mutate(cumul = cumsum(occ_date))

# Plot
ggplot(cumul_sp, aes(x = date, y = cumul, colour = datasource)) +
  geom_line() +
  theme_minimal()
```



# Map

You can also embed plots, for example:

```{r initial code for maps, echo=FALSE}


# Load the polygon
aut <- getData('GADM', country = 'aut', level = 0)
aut <- st_as_sf(aut)

med_basin

# Create the grid ((Try with polygons
grid_med = med_basin %>% 
  st_make_grid(cellsize = 0.3, # A general rule of thumb is that 1 degree of latitude ~ 111 kilometres. Here 0.3×1110 ~ 33 kilometres.
               what = "polygons") %>% 
  st_intersection(med_basin)    

plot(grid_med)

# Create a dataset for occurence
occ_spat_data = st_as_sf(data_plant_med, 
                         coords = c("decimallongitude", "decimallatitude"), 
                         crs = 4326) # WGS 84

# Intersect the dataset
tab = st_intersects(grid_med, occ_spat_data)
lengths(tab)

res_grid = st_sf(density = lengths(tab), 
                 geometry = st_cast(grid_med, "MULTIPOLYGON"))


# plot
palette_couleurs <- brewer.pal(6, "Blues") #
palette_couleurs = c('#f7fcfd','#e5f5f9','#ccece6','#99d8c9','#66c2a4','#41ae76','#238b45','#005824')

ggplot() +
  geom_sf(data = res_grid, aes(fill = density),lwd = 0) + # polygons filled based on the density value
  scale_fill_gradientn(colors = palette_couleurs) +
  # + # add the continent
  # + # add the sutyd area
  #scale_fill_viridis_c(option = "magma",begin = 0.1) +  # correct syntax for passing a hardcoded colour to a geom is to put it outside of the aes
  theme_minimal() +
  labs(title = "My graph")


# Dynamique view
# mapview(res_grid, zcol = "n")



#

# Créer le graphique avec ggplot2
ggplot(data, aes(x = x, y = y, fill = z)) +
  geom_tile() +
  
  labs(title = "Graphique avec palette de couleurs RColorBrewer")


```

## Spatial distribution occurences 

```{r}

```


## Spatial distribution of species richness
```{r}

```



## Complementarity of the datasets
```{r complementarity}

table(data_plant_med%>%filter(datasource == "inat")%>%select(species))

```

```{r}


#Trial with sf package 
# load some spatial data. Administrative Boundary
#https://stackoverflow.com/questions/41787313/how-to-create-a-grid-of-       spatial-points
# aut <- getData('GADM', country = 'aut', level = 0)
# aut <- st_as_sf(aut)
# #Try with polygons
# grid <- aut %>% 
#   st_make_grid(cellsize = 0.5, what = "polygons") %>% 
#   st_intersection(aut)                               

#fake data
lat<-runif(1000, 46.5, 48.5)
lon<-runif(1000, 13,16)
pos<-data.frame(lat,lon)

pos = st_as_sf(pos, coords = c("lon", "lat"), crs = 4326)

tab = st_intersects(grid, pos)
lengths(tab)

grid = st_sf(n = lengths(tab), geometry = st_cast(grid, "MULTIPOLYGON"))

mapview(grid, zcol = "n")


####################"


library(ggplot2)
library(dplyr)

# Créer des données de plantes (exemple)
set.seed(123)
plantes <- data.frame(
  longitude = runif(100, -180, 180),
  latitude = runif(100, -90, 90)
)

# Créer une grille raster
resolution <- 2  # Définir la résolution de la grille en degrés
lon_bins <- seq(-180, 180, by = resolution)
lat_bins <- seq(-90, 90, by = resolution)

grille_raster <- expand.grid(longitude = lon_bins, latitude = lat_bins)

# Compter les observations par cellule
observations_par_cellule <- plantes %>%
  mutate(
    cellule_lon = cut(longitude, breaks = lon_bins, labels = FALSE),
    cellule_lat = cut(latitude, breaks = lat_bins, labels = FALSE)
  ) %>%
  group_by(cellule_lon, cellule_lat) %>%
  summarise(nombre_observations = n(), .groups = "drop")

# Fusionner la grille avec les données d'observations
grille_raster <- left_join(grille_raster, observations_par_cellule, 
                           by = c("longitude" = "cellule_lon", "latitude" = "cellule_lat"))

# Créer la carte avec ggplot2
ggplot(grille_raster, aes(x = longitude, y = latitude, fill = nombre_observations)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "blue", na.value = "grey50", name = "Observations") +
  labs(title = "Nombre d'observations de plantes par cellule")



```

